The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-09-29 Paper/jaccardbaetica.tex
correction finished

Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}. Morisita-Horn index gives a value between 0 (sites have an exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes). To apply Morisita-Horn we firstly calculated the number of times that each stamp code appeared in an amphora workshop. 
This method allowed us to bear in mind the number of frequency of stamps for each workshop. 
%This method allowed us to bear in mind a similar number of times for each repeated stamp per workshop\xavi{no entiendo esta frase}.

@@

The lower similarity of stamp codes in \textit{Britannia} compared to \textit{Baetica} could be explained by the spatial correlation as the area under study is much bigger and the sites are more distant between them. This can be observed more clearly on military sites as the ones with higher similarity are geographically closer. Two patterns can be observed on the dendrogram: 1) nearby sites tend to share more similar stamps and 2) a majority of stamp code are only found on one site.

@@

A higher concentration of sites sharing similar stamps were found in areas eminently militarised and close to German \textit{limes}, although most sites mostly showed different stamps. 
\maria{he cambiado esta frase}
%even if most sites mostly showed different stamps~\xavi{quieres decir que los sitios más cercanos al limes son más similares? no queda claro y hay que detallarlo un poco más}.

@@

This work has explored to what extent the distribution of amphoric stamps can provide insights on the organisation of olive oil Roman markets both at the production and consumption areas. For this reason, an index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres.

@@

The analysis of the amphora workshops in \textit{Baetica} province suggests that a majority of stamp codes were only used on a single workshop. Beyond this clear pattern it seems that the similarity of stamp codes is correlated with spatial distance and the amphoric production of spatially close sites shared some stamp codes. Particularly relevant here is the fact that our results show how similar stamp codes were found on sites belonging to the same \textit{conventus} area. This result could indicate that the workshops of the different \textit{conventus} used different stamp codes that were not shared between them. Despite these stamps tend to share the same area of production, we do not identify groups of any more than three workshops sharing the same amphoric stamps as the dendrogram showed.

@@

Both consumption areas showed a correlation between spatial distance of sites and their similarity of amphoric stamps. In the case of  \textit{Britannia}, the correlation was higher than \textit{Germania}. In \textit{Britannia} the majority of similarity stamps were mainly found in military centres.

@@

This work has used quantitative analysis to explore long-range trade of olive oil using the stamps found in Dressel 20 vessels. The most interesting result is linked to the meaning of these inscriptions. The use of specific codes seems to be almost exclusively decided by each workshop, so it could be used to identify amphorae found in consumption centers (i.e. an amphorae with a given stamp was probably made on a specific workshop). This idea could explain why the similarity of stamps between workshops is generally very low. On the other hand, some code stamps have been identified in nearby workshops so this pattern suggests that the clustered workshops were somehow connected (e.g. they belonged to the same owner).

@@

This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros, Víctor M. Lozano and Ignacio Morer for their useful suggestions.  
All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}. Source and code are available at \maria{incluir github}.

>>>>

Considering our dataset as a non-uniform sample, this method provides a useful tool to handle large samples with different sizes and diversity \citep{wolda_similarity_1981}. Morisita-Horn index gives a value between 0 (sites have exact stamp codes and frequency distribution) and 1 (complete difference between stamp codes). To apply Morisita-Horn we first calculated the number of times that each stamp code appeared in an amphora workshop. 
This method allowed us to bear in mind the number of frequency of stamps for each workshop. 
%This method allowed us to bear in mind a similar number of times for each repeated stamp per workshop\xavi{no entiendo esta frase}.

@@

The lower similarity of stamp codes in \textit{Britannia} compared to \textit{Baetica} could be explained by the spatial correlation as the area under study is much bigger and the sites are more distant between them. This can be observed more clearly on military sites as the ones with higher similarity are geographically closer. Two patterns can be observed on the dendrogram: 1) nearby sites tend to share more similar stamps and 2) a majority of stamp codes are only found in one site.

@@

A higher concentration of sites sharing similar stamps was found in areas eminently militarised and close to German \textit{limes}, although most sites mostly showed different stamps. 
\maria{he cambiado esta frase}
%even if most sites mostly showed different stamps~\xavi{quieres decir que los sitios más cercanos al limes son más similares? no queda claro y hay que detallarlo un poco más}.

@@

This work has explored to what extent the distribution of amphoric stamps can provide insights into the organisation of olive oil Roman markets both at the production and consumption areas. For this reason, an index of dissimilarity was used to detect differences between the distribution of the amphoric stamps and the spatial distance of producers and consumption centres.

@@

The analysis of the amphora workshops in \textit{Baetica} province suggests that a majority of stamp codes were only used on a single workshop. Beyond this clear pattern, it seems that the similarity of stamp codes is correlated with spatial distance and the amphoric production of spatially close sites shared some stamp codes. Particularly relevant here is the fact that our results show how similar stamp codes were found on sites belonging to the same \textit{conventus} area. This result could indicate that the workshops of the different \textit{conventus} used different stamp codes that were not shared between them. Despite these stamps tend to share the same area of production, we do not identify groups of any more than three workshops sharing the same amphoric stamps as the dendrogram showed.

@@

Both consumption areas showed a correlation between the spatial distance of sites and their similarity of amphoric stamps. In the case of  \textit{Britannia}, the correlation was higher than \textit{Germania}. In \textit{Britannia} the majority of similarity stamps were mainly found in military centres.

@@

This work has used quantitative analysis to explore long-range trade of olive oil using the stamps found in Dressel 20 vessels. The most interesting result is linked to the meaning of these inscriptions. The use of specific codes seems to be almost exclusively decided by each workshop, so it could be used to identify amphorae found in consumption centers (i.e. an amphora with a given stamp was probably made on a specific workshop). This idea could explain why the similarity of stamps between workshops is generally very low. On the other hand, some code stamps have been identified in nearby workshops so this pattern suggests that the clustered workshops were somehow connected (e.g. they belonged to the same owner).

@@

This research was partly funded by the European Research Council Advanced Grant EPNet (340828). We are grateful to Simon Carrignon, Juan Moros and Ignacio Morer for their useful suggestions.  
All data has been analysed and conducted in R program version 3.2.4, using the packages \textit{vegan} \citep{oksanen_vegan_2007}, \textit{ggplot2} \citep{ggplot2:_2016}. Source and code are available at \maria{incluir github}.
